We performed post-attendance surveys for AHW, GHW and NHW in 2016 and 2017. All surveys contained general questions about attitudes towards the workshop as well as open science and reproducibility, and their own skills in statistical and computational methods, shared among all three surveys. The NHW and GHW surveys and the AHW 2017 survey were administered on site on the last day of the workshop; AHW participants in 2016 were e-mailed immediately after the workshop, with reminders after several weeks. All responses were collected within four weeks of the end of the workshop. The experimental procedures were approved by the Institutional Review Boards at UW, NYU and UC Berkeley. All participants gave their informed consent.

Response rates for NHW (2016: 41 responses; 2017: 45 responses) and GHW (2016: 42 responses; 2017: 41 responses) were 100\% in both years; the response rate for AHW was 71\% (35 out of 49) in 2016 and 82\% (37 out of 45) in 2017. The lower response rates for AHW can be explained by the generally lower rates of response for time-delayed survey in 2016. In 2017, a number of participants did not attend the full week, and thus several attendees were not present on the last day when the survey was administered. For the AHW surveys, we checked whether the demographics of the respondents different significantly from those of the attendees overall in three relevant categories (career stage, racial/ethnic background and gender identity) and find no significant difference in any of the categories except for career stage in 2017: here, we find that graduate students are underrepresented (27\% of respondents compared to 41\% of attendees), while undergraduate students are overrepresented (25\% of respondents compared to 12\% of attendees). Because both undergraduate and graduate students are at the beginning of their careers for whom AHW is likely the first exposure of many of the concepts around data science and open science, we believe that this discrepancy has little impact on the overall results presented here.

Participants were asked to respond to statements regarding these topics using a six-point Likert-type scale. All questions were anonymously recorded. No responses were discarded, and no pre-processing was performed on the data. We test for correlations between categorical variables using a standard $\chi^2$ test and compute the effect sizes via a bias-corrected version of Cramer's V \cite{bergsma2013}. Only $p$-values with $p < 0.0018$ (equivalent to $p < 0.05$ corrected for $N=27$ trials) were reported as potential candidates. The full analysis procedure is available online\footnote{See the repository: \url{https://github.com/uwescience/HackWeek-Writeup}}.
